---
title: "An exploration of the world's wines"
author: "Andy Pagès"
date: "2018-03-20"
output: pdf_document
---

# Introduction 

Wine is a famous alcoholic drink made from fermented grapes that has been produced for thousands of years. The process to create wine 
is very complex and vary from one place of the world to another. Using a data set connecting wines with their price, origin, location, designation,
variety, reviews and more, this data exploration aims to first discover as a consumer where to find the best wines, how is the quality affected by location and 
winery, but also to dig deeper into what makes up a good wine and what are the present correlations to better understand the fascinating world of wine production.

The first initial questions that we will try to answer are:

1. How does the quality of wine change according to its origin country?
2. Which are the best wines according to quality and price? Where are they located?
3. What words and designations are usually used to describe the best wines? 

# Setup

## Libraries

The libraries used in this project are:

- *futile.logger*: A logging library for clearer output,
- *ggplot2*: The famous plotting library,
- *dplyr*: The famour data manipulation library,
- *maps* and *scales*: Used for map creations.

```{r}

# Packages
library("futile.logger")
library("ggplot2")
library("dplyr", warn.conflicts = FALSE)
library("scales")
library("maps")

```

## Modules

The project structure is split in 3 types of modules:

- *plots*: Includes generic functions to print specific types of plots,
- *questions*: Includes functions that contain the code for each question,
- *utils*: Includes generic functions to perform common tasks on a dataset.

```{r}

# Plots
source (file="./plots/histogram.R")
source (file="./plots/map.R")
source (file="./plots/pie.R")
source (file="./plots/box.R")

# Questions
source (file="./questions/q1.R")

# Utils
source (file="./utils/clean.R")

```

# Exploration

## Basics

Let's first import the data from the *CSV* file, then utilize R's basic tools to get a first glance at the data. We will also hide
the warnings for better readability.

```{r}

options(warn=-1)

# Loading data
flog.info('Loading data..')
if (!exists("data.raw")) {
    data.raw = read.csv("data_sets/wines/winemag-150k.csv")
}

# Info
flog.info("%d rows loaded", nrow(data.raw))

# Cleaning data
data.clean = clean(data.raw)

# Summary
summary(data.clean)

```

This summary gives us basic information on this data set. A relevant element to note is that the points system used
to rate the wines ranges from 80 to 100. We can therefore assume that relatively, a "*bad*" wine is given 80 points and a "*good*" wine is given 100 points.
The prices on the other hand range from $4 to $2300, but the 3rd quartile indicates that only 25% of the prices are above $40. Finally, general insights for
the countries and regions can be extracted, such as the United States being the most active wine producer with almost twice France's or Italy's production numbers.

\newpage

## Q1: How does the quality of wine change according to its origin country?

Let's first take a look at the general production of wine world-wide. What are the top 10 countries producing the most wines? (see **./plots/histogram.R** 
to see the code used to plot).

```{r eval=FALSE}
q1_best_10_production = function(dataset) {
    data = dataset %>% 
                group_by(country) %>% 
                summarise(count = n()) %>% 
                arrange(desc(count)) %>%
                head(n = 10)
    plotHistogram(data, "country", "count", "country", 
        "Top 10 countries", "Wine producers", sorted = TRUE)
}
```

```{r fig.height = 3, fig.width = 7, fig.align = "center"}

q1_best_10_production(data.clean)

```

From this histogram, it is clear that the United States are leader in terms of numbers of wine produced. This is to be expected
from the size of the United States compared to countries like Italy or France. However, quantity does not imply quality. Let's
get an overview of the 10 best countries in term of wine quality, using a mean of their scores. We will here only consider the top 10 countries 
in term of production (> 3000 different wines):

```{r eval=FALSE}
q1_best_10_quality = function(dataset) {
    data = dataset %>% 
                group_by(country) %>% 
                summarise(count = n(), average = mean(points)) %>%
                filter(count >= 3000) %>%
                arrange(desc(average)) %>%
                head(n = 10)
    plotHistogram(data, "country", "average", "country", "Top 10 countries", "Wine quality", 
        sorted = TRUE, ylim = c(min(data$average), max(data$average)))
}
```

```{r fig.height = 3, fig.width = 7, fig.align = "center"}

q1_best_10_quality(data.clean)

```

The leading trio is hence Austria, France and Italy. Before taking a deeper look into the leading countries quality-wise and understand
exactly how their wine quality differ, let's take a more world-wide and visual view at the quality of wine using their mean scores 
with a minimum of 10 produced wines (see **./plots/map.R** to see the code used to plot):

```{r eval=FALSE}
q1_worldwide_quality = function(dataset) {
    data = dataset %>% 
                group_by(country) %>% 
                summarise(count = n(), average = mean(points)) %>%
                filter(count >= 10)
    plotMap(data, "country", "average", "World-wide wine quality", *
        "Based on rankings from 80 to 100")
    plotMap(data, "country", "average", "Europe wine quality", 
        "Based on rankings from 80 to 100", type = "europe")
}
```

```{r fig.height = 4.4}

q1_worldwide_quality(data.clean)

```

Let's now take a deeper look at the top 10 countries quality-wise and understand exactly how they differ and which 
of those is actually the "best wine country".

```{r}
countries.best = data.clean %>%
                group_by(country) %>% 
                summarise(count = n(), average = mean(points)) %>%
                filter(count >= 3000) %>%
                arrange(desc(average)) %>%
                head(n = 10) %>%
                pull(country)
data.best = data.clean %>% filter(country %in% countries.best)
```

In this 10 countries, the number of wines produced for each can highly vary. This variation in production
can biase the average and hence must be considered to determine the best wine countries. We can first
take a look at the repartition of the total produced wine from these 10 countries:

```{r eval=FALSE}
q1_best_10_production_repartition = function(dataset) {
    total = nrow(dataset)
    data = dataset %>%
                group_by(country) %>% 
                summarise(count = n(), percentage = count / total)
    plotPie(data, "country", "count", "Production repartition", "Top 10 wine countries")
}
```

```{r fig.height = 3, fig.width = 7, fig.align = "center"}

q1_best_10_production_repartition(data.best)

```

This pie chart shows how the US, France and Italy dominates in terms of production numbers. Comparing
these numbers raises questions regarding the general quality of a country. For example, Austria's wine quality mean
is superior to France's, but it is clear here that France produces a lot more wine than Austria, which country is actually better?
This will require to look at how the data spreads, the quartiles and variations of each country.

```{r eval=FALSE}
q1_best_10_data_spread = function(dataset) {
    plotBox(dataset, "country", "points", "Top 10 countries", "Data variations")
}
```

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
q1_best_10_data_spread(data.best)
```

Many insights can be taken from this graph. Let's focus on the three best candidates to be the best wine country: Austria, France, Italy and US.
These three countries have very close data, especially Austria and France with almost identical medians and very close means. 
We can first however notice that France, with a lot more wines, still present a range box just a little bigger than Austria's. That indicates
that even though France has a lot more wines, their scores are narrowly centered around it's median and mean, which implies a good quality for 
so many different wines. France and Italy also present more outliers, especially with Italy that shows a very concentrated data spread (however lower) 
with some very good wines and a bit less of very bad wines compare to its mean. The United State, with a lower mean and median,
still present very good wines. Its range box is however wider than the other three top countries, which indicates a less consistent
wine quality. While this graph gives interesting insights regarding the spread of the data, it is not enough to distinguish the best wine country.

To conclude on the variation of wine quality according to its origin, let's take a look at the 30 best wines in the world and their origin:

```{r eval=FALSE}
q1_best_30_wines = function(dataset) {
    wines.best = dataset %>%
                arrange(desc(points)) %>%
                head(n = 30) %>%
                group_by(country) %>%
                summarise(count = n())
    plotHistogram(wines.best, "country", "count", "country", 
        "Top 30 wines", "Countries repartition", sorted = TRUE)
}
```

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
q1_best_30_wines(data.clean)
```

This shows that the United States seem to produce the largest amount of excellent wines. Austria is here discarded from the top
countries. The past results have shown that countries present many complex factors influencing wine quality, making
it hard to distinguish one absolute leader in terms of quality. The term of "best wine"
is of course subjective and might be influenced also by the price, but here we focused only on the scores. 
If distinguishing countries appears like unconsistent research, then let's focus on a more narrow study, introducing a new question: 
**Question 4: Which are the best wineries in the world?**