---
title: "An exploration of the world's wines"
author: "Andy Pagès"
date: '2018-03-20'
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Introduction 

Wine is a famous alcoholic drink made from fermented grapes that has been produced for thousands of years. The process to create wine 
is very complex and vary from one place of the world to another. Using a data set connecting wines with their price, origin, location, designation,
variety, reviews and more, this data exploration aims to first discover as a consumer where to find the best wines, how is the quality affected by location and 
winery, but also to dig deeper into what makes up a good wine and what are the present correlations to better understand the fascinating world of wine production.

The first initial questions that we will try to answer are:

1. How does the quality of wine change according to its origin country?
2. What is the relation between quality and price? 
3. What words and designations are usually used to describe the best wines? 

# Setup

## Libraries

The libraries used in this project are:

```{r}
# Packages
library("futile.logger")
library("ggplot2")
library("dplyr", warn.conflicts = FALSE)
library("scales")
library("maps")
library("tm")
library("SnowballC")
library("wordcloud")
library("RColorBrewer")
```

## Modules

The project structure is split in 2 types of modules:

- *plots*: Includes generic functions to print specific types of plots,
- *utils*: Includes generic functions to perform common tasks on a dataset.

```{r}

# Plots
source (file="./plots/histogram.R")
source (file="./plots/map.R")
source (file="./plots/pie.R")
source (file="./plots/box.R")
source (file="./plots/point.R")

# Utils
source (file="./utils/clean.R")
source (file="./utils/text.R")

```

# Exploration

## Basics

Let's first import the data from the *CSV* file, then utilize R's basic tools to get a first glance at the data. We will also hide
the warnings for better readability.

```{r}

options(warn=-1)

# Loading data
flog.info('Loading data..')
if (!exists("data.raw")) {
    data.raw = read.csv("data_sets/wines/winemag-150k.csv")
}

# Info
flog.info("%d rows loaded", nrow(data.raw))

# Cleaning data
data.clean = clean(data.raw)

# Summary
summary(data.clean)

```

This summary gives us basic information on this data set. A relevant element to note is that the points system used
to rate the wines ranges from 80 to 100. We can therefore assume that relatively, a "*worst*" wine is given 80 points and a "*best*" wine is given 100 points.
The prices on the other hand range from $4 to $2300, but the 3rd quartile indicates that only 25% of the prices are above $40. Finally, general insights for
the countries and regions can be extracted, such as the United States being the most active wine producer with almost twice France's or Italy's production numbers.

\newpage

## Q1: How does the quality of wine change according to its origin country?

Let's first get an overview of the 10 best countries in term of wine quality, using a mean of their scores. 

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data = data.clean %>% 
            group_by(country) %>% 
            summarise(count = n(), average = mean(points)) %>%
            filter(count >= 3000) %>%
            arrange(desc(average)) %>%
            head(n = 10)
plotHistogram(data, "country", "average", "country", "Top 10 countries", "Wine quality", 
    sorted = TRUE, ylim = c(min(data$average), max(data$average)))
```

The leading trio is hence France and Italy and Portugal, with France leading. Before taking a deeper look into the leading countries quality-wise and understand
exactly how their wine quality differ, let's take a more world-wide and visual view at the quality of wine using their mean scores 
with a minimum of 10 produced wines (see **./plots/map.R** to see the code used to plot):

```{r fig.height = 4.2}
data = data.clean %>% 
            group_by(country) %>% 
            summarise(count = n(), average = mean(points)) %>%
            filter(count >= 10)
plotMap(data, "country", "average", "World-wide wine quality", 
    "Based on rankings from 80 to 100")
plotMap(data, "country", "average", "Europe wine quality", 
    "Based on rankings from 80 to 100", type = "europe")
```

Let's now take a deeper look at the top 10 countries quality-wise (considering countries that have produced at least 3000 wines)
 and understand exactly how they differ and which of those is actually the "best wine country".

```{r}
countries.best = data.clean %>%
                group_by(country) %>% 
                summarise(count = n(), average = mean(points)) %>%
                filter(count >= 3000) %>%
                arrange(desc(average)) %>%
                head(n = 10) %>%
                pull(country)
data.best = data.clean %>% filter(country %in% countries.best)
```

In this 10 countries, the number of wines produced for each can highly vary. This variation in production
can biase the average and hence must be considered to determine the best wine countries. We can first
take a look at the repartition of the total produced wine from these 10 countries:


```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data = data.best %>%
            group_by(country) %>% 
            summarise(count = n())
plotPie(data, "country", "count", "Production repartition", "Top 10 wine countries")
```

This pie chart shows how the US, France and Italy dominates in terms of production numbers. Comparing
these numbers raises questions regarding the general quality of a country. For example, Portugals's wine quality mean
is superior to US's, but it is clear here that the US produce a lot more wine than Portugal, which country is actually better?
This will require to look at how the data spreads, the quartiles and variations of each country.

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
plotBox(data.best, "country", "points", "Top 10 countries", "Data variations")
```

Many insights can be taken from this graph. These top countries have very close data, with very close means and medians. 
We can first however notice that France present a range bigger than Italy's. That indicates than even though France's average quality is superior,
Italy has a more concentrated data spread, which means it has a more consistent wine quality. France however, with a less skewed median,
shows a more even distribution of good and bad wines. The United State, with a lower mean and median,
still present very good wines. Its range box is however wider than the other three top countries, which indicates a less consistent
wine quality. While this graph gives interesting insights regarding the spread of the data, it is not enough to distinguish the best wine country.
But the leading trio is France, Italy and the US.

To conclude on the variation of wine quality according to its origin, let's take a look at the 30 best wines in the world and their origin:

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data = data.clean %>%
            arrange(desc(points)) %>%
            head(n = 30) %>%
            group_by(country) %>%
            summarise(count = n())
plotHistogram(data, "country", "count", "country", "Top 30 wines", "Countries repartition", sorted = TRUE)
```

This shows that the United States seem to produce the largest amount of excellent wines. The past results have shown that countries present many 
complex factors influencing wine quality, making it hard to distinguish one absolute leader in terms of quality. The term of "best wine"
is of course subjective and might be influenced also by the price, but here we focused only on the scores. 
If distinguishing countries appears like inconsistent research, then let's focus on a more narrow study, introducing a new question: 
**Question 4: Which are the best wineries in the world?**


## Q2: What is the relation between quality and price? 

From the first question, we can assume that a "very good wine" is a wine with a score superior or equal to 95. 
However the price of a wine is an important matter. Let's start by getting an overview of the prices, the average
price, the lowest and cheapest, etc.

First, because the dataset contains wines with no price, let's clean the data by removing these rows:

```{r}
data.prices = removeNa(data.clean, "price")
```

Let's take a look at a summary:

```{r}
summary(data.prices$price)
```

This summary shows that the average wine costs $33.66. The third quartile being only $40
indicates that high priced wines (above $100) are a very small portion of the data. To get
a first, more general and "down to earth" view of the wines prices variations, let's first focus
on wines with a price inferior or equal to $100.

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
p_mean = mean(data.prices$price)
data.prices.noutlier = data.prices %>% filter(price <= 100)
ggplot(data.prices.noutlier, aes(x = price)) +
    geom_histogram(binwidth = 10, color = "#efa352", fill = "#efe553") +
    geom_vline(aes(xintercept = p_mean), color = "#444444", linetype = "dashed", size = 1) +
    labs(title = "Prices distribution", subtitle = "Wines under $100") +
    theme_bw() +
    scale_x_continuous(breaks = seq(0, 100, 10)) 
```

We can see that most prices are between $10 and $30, and a smaller portion a more expensive with a range of $30 - $50.
Before taking a specific look at the best wines and which of them have the best prices for their quality, let's look at how
prices evolve according to quality. For a better visualization, the extreme outliers (> $1000) are not included here.

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
plotPoint(data.prices, "points", "price", "Evolution of price according to quality", 
    "Wines under $1000", ylim = c(0, 1000))
```

This graph clearly shows how the price of wines from scores 80 - 90 (bad to middle quality wines) slighty increase. However,
when wines start to approach scores close to 95, the prices significantly increase faster.

Let's now focus on the the prices of the best wines (> 95 points). 

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.prices.best = data.prices %>% filter(points >= 95)
plotBox(data.prices.best, "points", "price", "Price distribution", 
    "Wines with a quality ranked in 95 - 100")
```

This graph shows how the prices are a lot more concentrated for wines of lesser quality (95 - 97) with outliers
that also stay under a certain range. On the other hand, the best wines (98 - 100) have a less regular pricing
and extreme outliers ($2300 for a 99 wine). It shows how these wines have a more "random" and "unjustified" pricing depending
on special wines and occasions.

Let's now take a deeper look at quartiles, again for each quality category from 95 to 100, and how they are distributed.
We can assume the following: Q1 are cheap wines, Q2 are average prices wines, Q3 are expensive wines and Q4 are very expensive wines.

```{r fig.height = 5, fig.width = 7, fig.align = "center"}
cheap = data.prices.best %>% 
        group_by(points) %>% 
        summarise(price_category = "1: cheap", count = sum(price <= quantile(price)[2]))
average = data.prices.best %>% 
        group_by(points) %>% 
        summarise(price_category = "2: average", 
            count = sum(price > quantile(price)[2] & price <= quantile(price)[3]))
expensive = data.prices.best %>% 
            group_by(points) %>% 
            summarise(price_category = "3: expensive", 
                count = sum(price > quantile(price)[3] & price <= quantile(price)[4]))
very_expensive = data.prices.best %>% 
                group_by(points) %>% 
                summarise(price_category = "4: very_expensive", 
                    count = sum(price > quantile(price)[4]))
data.prices.categories = rbind(cheap, average, expensive, very_expensive)

h = histogram(data.prices.categories, "price_category", "count", "price_category", 
        "Price categories", "For each quality range (95 - 100)", legend = TRUE) + 
    facet_wrap(~ points, scale = "free") +
    scale_fill_manual(
        name = "Category", 
        values = c("#6AA84F", "#4D9DE0", "#E15554", "#E1BC29"), 
        labels = c("Cheap", "Average", "Expensive", "Very expensive")) +
    xlab("Price category") +
    theme(
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank())
plot(h)
```

This plot shows that wines from the quality category 96 have a very good price distribution, where
a decreasing number from cheap to very expensive. Qualities 95 to 97 are more accessible with a proportion 
of cheap wines dominating, whereas qualities 98 - 100 seem to present more very expensive wines, especially 
for wines of quality 99 and 100.

With the fact that wines of quality 96 are a very interesting subset of the "best wines", a new question emerges: **Q5: Where can 
be found the best wines for the best prices ? (Quality 96)**

## Q3: What words and designations are usually used to describe the best wines?

Each wine is described a few sentences. Let's use this and do some text mining to see what words are usually
used to describe good wines.

```{r fig.height = 5, fig.width = 7, fig.align = "center"}
data.words = data.clean %>% filter(points >= 95)
data.words.freq = get_words_frequencies(data.words$description, 
    c("wine", "drink", "years", "now", "finish", "will", "shows", 
    "vineyard", "just", "yet", "well", "fine", "flavors", "one", "notes", "made",
    "best", "also", "even", "great", "like", "next", "mouth", 
    "almost", "new", "fruit", "palate", "aromas"))
data.words.freq = data.words.freq %>% filter(freq != 312)
wordcloud(data.words.freq$word, data.words.freq$freq, min.freq = 1,
          max.words=50, colors = brewer.pal(8, "Paired"))

```

Let's plot have a more precise view on what are the 10 words that are the most used to describe good wines:

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.freq.top10 = data.words.freq %>% head(n = 10)
plotHistogram(data.freq.top10, "word", "freq", "word", 
    "Top 10 description words", "For wines of a quality in 95 - 100", sorted = TRUE)
```

Those are therefore the words to look for when searching for a great wine. Let's take a quick look at the words
that describe "bad" wines, that is wines with a quality equal or under 85:

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.words = data.clean %>% filter(points <= 85)
data.words.freq = get_words_frequencies(data.words$description, 
    c("wine", "drink", "years", "now", "finish", "will", "shows", 
    "vineyard", "just", "yet", "well", "fine", "flavors", "one", "notes", "made",
    "best", "also", "even", "great", "like", "next", "mouth", 
    "almost", "new", "fruit", "palate", "aromas"))
data.words.freq = data.words.freq %>% filter(freq != 312)
data.freq.top10 = data.words.freq %>% head(n = 10)
plotHistogram(data.freq.top10, "word", "freq", "word", 
    "Top 10 description words", "For wines of a quality in 80 - 85", sorted = TRUE)
```

We can see that good wines are usually described with words like "tannins", "black", "ripe" or "oak", whereas 
lesser wines are described with "cherry", "sweet", "dry" or "soft". This makes sense as very good and expensive
wines are usually targeted at connaisseurs that look for very specific and rare tastes, whereas cheaper and more common
wines are targeted to a more casual population, looking for simple and sweeter tastes.

\newpage

## Q4: Question 4: Which are the best wineries in the world?

Let's first take a look a the 10 best wineries based on their average points, considering
that they have produced at least 10 wines:

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.wineries = data.clean %>% 
              group_by(winery, country) %>% 
              summarise(count = n(), mean = mean(points)) %>% 
              filter(count > 10) %>%
              arrange(desc(mean))
knitr::kable(head(data.wineries, 10), format = "markdown")
```

We can see here that France dominates with a majority of wineries present in the top 10. Of course,
the US still seem to be bigger producers of wine. If we restrict the minimum number of produced wines to 50, we can cleary see that
the US dominate:

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.wineries = data.clean %>% 
              group_by(winery, country) %>% 
              summarise(count = n(), mean = mean(points)) %>% 
              filter(count > 50) %>%
              arrange(desc(mean))
knitr::kable(head(data.wineries, 10), format = "markdown")
```

Let's take a look a the distribution of the number of wines produced per wineries to get an idea
of a minimum number to consider for the best winery:

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.wineries = data.clean %>% 
              group_by(winery) %>% 
              summarise(produced = n()) %>%
              filter(produced >= 10 & produced <= 100)
ggplot(data.wineries, aes(x = produced)) + 
  geom_histogram(binwidth = 5, color = "#efa352", fill = "#efe553") +
  theme_bw() +
  scale_x_continuous(breaks = seq(0, 100, 5))
```

This shows that most wineries produce around 10 and 30 wines.
Because France, Italy and US again seem to be the leaders, let's extract for each of these countries the best
winery with a minimum of 10 produces wines:


```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.wineries = data.clean %>% 
              group_by(winery, country) %>% 
              summarise(count = n(), mean = mean(points)) %>% 
              filter(count > 50)
data.wineries.france = data.wineries %>%
                      filter(country == "France") %>%
                      arrange(desc(mean)) %>%
                      head(1)
data.wineries.italy = data.wineries %>%
                      filter(country == "Italy") %>%
                      arrange(desc(mean)) %>%
                      head(1)
data.wineries.us = data.wineries %>%
                      filter(country == "US") %>%
                      arrange(desc(mean)) %>%
                      head(1)
data.wineries.best = rbind(data.wineries.france, data.wineries.italy, data.wineries.us) %>%
                    arrange(desc(mean))
knitr::kable(data.wineries.best, format = "markdown")

```

These three wineries can hence be considered the best wineries in the word. Let's take a closer look at these wineries
by comparing the varieties of wine they offer:

```{r fig.height = 5, fig.width = 7, fig.align = "center"}
wineries = data.wineries.best %>% pull(winery)
data.wineries = data.clean %>% filter(winery %in% wineries)
data.varieties = data.wineries %>%
                group_by(winery, variety) %>%
                summarise(count = n())
h = histogram(data.varieties, "variety", "count", "winery",
    "Most common varieties", "For the three best wineries", legend = TRUE, sorted = TRUE) +
  coord_flip() + 
  scale_fill_discrete(name = "Winery")
plot(h)
```

## Q5: Where can be found the best wines for the best prices? (Quality 96)

Knowing that wines of quality 96 represent the best category to find avantageous prices, let's plot on the world map
the locations of such wines:

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.96 = data.clean %>% filter(!is.na(price)) %>% filter(points == 96)
data.96.q1 = quantile(data.96$price)[2]
data.96 = data.96 %>% filter(price <= data.96.q1)
data.96.countries = data.96 %>%
              group_by(country) %>%
              summarise(count = n())
plotPie(data.96.countries, "country", "count", "Amount of best/cheap wines", 
        "Based on wines of quality 96")
```

It is hence obvious that the US are the place to find very good and cheap wines. Let's take a closer look at the US and see where
exactly these advantageous wines can be found: 

```{r fig.height = 3, fig.width = 7, fig.align = "center"}
data.96.us = data.96 %>% filter(country == "US")
data.96.us = data.96.us %>%
            group_by(province) %>%
            summarise(count = n()) %>%
            mutate(province = tolower(province))
plotMap(data.96.us, "province", "count", "USA's best/cheap wines repartition", 
        "Based on wines of quality 96", type = "usa")
```

This map shows that the best place to find good wines at accessible price is the West coast of the United States, California in particular.

## Conclusion

During this data exploration we learned that:

- France is the country producing the best wines in terms of quality, followed by Italy and Portugal,
- The United States are the biggest producers of wine,
- South America countries produce bad to average quality wines,
- The leading trio in terms of quality and production numbers are France, Italy and the United States,
- The United States produce more of the world's best wines,
- The average price of a wine is $33.66,
- Most wines are priced between \$10 and \$50,
- The price increases alongside quality but higher quality wines (> 95) have a stronger price increase, reaching prices
over $1000,
- Higher quality wines have a less concentrated and regular price spread,
- Wines ranked of quality 96 are the best candidates to find great wines at advantageous prices,
- Good and expensive wines are described with words such as tannins, black, ripe, cabernet or oak,
- Lesser and cheaper wines are described with words such as cherry, sweet, soft, dry and simple,
- The best wineries (producing at least 10 wines) are located in France, Italy and the United States,
- The best three wineries are Failla (US), Domaine Zind-Humbrecht (France) and Castello Banfi (Italy),
- The most common varities found in the best wineries are Pinot Noir, Gewürztraminer, Pinot Gris, Sangiovese or Riesling,
- The United States are the best place to find great wine at accessible price, especially in the West coast, California.

